Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-07-30 analysis/paper/hotspots-paper.Rmd
added Chr2 hotspot results

>>>>

The scatter plot of pleiotropy test statistics and mediation LOD differences reveals that few local-nonlocal 
trait pairs have both a high value of the pleiotropy test statistic and a high value of the mediation LOD 
difference. Many points have low values of both, while some points demonstrate a high value of only mediation 
LOD difference or pleiotropy test statistic.

Many points have negative values of mediation LOD difference. These correspond to pairs that, when conditioning 
on the putative mediator, the LOD for the nonlocal trait increased.

The points are color-coded to identify those pairs that involve the local trait *Hnf4a*, which 
@keller2018genetic identified as a key mediator of traits at this hotspot. Blue points involve *Hnf4a*, while 
red points do not. We see that many blue point have large mediation LOD differences and small pleiotropy test 
statistics. In fact, most of the points with mediation LOD difference above 15 involve *Hnf4a* and have small 
pleiotropy test statistics.

### 2019-07-30 analysis/paper/methods.Rmd
added Chr2 hotspot results

We created heatmaps to examine patterns among pleiotropy test results in each hotspot.
We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
trait and each column was an expression trait. Each cell, then, contained the pleiotropy test statistic
from the analysis involving the row's expression trait and the column's expression trait.
To create a heatmap, we first performed two hierarchical clusterings, one on the submatrix involving only
nonlocal expression traits and a second on the submatrix involving only local expression traits. We
calculated the Euclidean distances between all trait pairs before clustering. We ordered the rows and
columns according to optimal leaf ordering. We then combined the
local and nonlocal traits into a single heatmap. To achieve greater coloring contrasts in the heatmaps, we
set the color spectrum limits at 0 (dark blue) and 5 (yellow). Test statistics above 5 are represented by
yellow.

>>>>

### Heatmaps

We created heatmaps to examine patterns among pleiotropy test results in each hotspot.
We represented the pleiotropy test statistics as a symmetric matrix, where each row was one expression
trait and each column was an expression trait. 
Excepting the cells on the diagonal, each cell, then, contained the pleiotropy test statistic
from the analysis involving the row's expression trait and the column's expression trait.
To create a heatmap, we first performed two hierarchical clusterings of rows. 
One hierarchical clustering involved only the rows and columns containing local traits, while the other contained only 
rows and columns containing nonlocal traits.
We calculated the Euclidean distances between all row pairs (in the respective submatrices) before clustering. We ordered the rows according to optimal leaf ordering [@hahsler2008getting;@dendextend]. 
We then combined the
local and nonlocal traits into a single heatmap for each hotspot. 
To achieve coloring contrasts in the heatmaps, we
set the color spectrum limits at 0 (dark blue) and 5 (yellow). Pleiotropy test statistics above 5, too, are 
represented by
yellow. We annotated rows by a collection of columns on the right-hand side of the heatmap. This binary matrix 
consists of columns that summarize mediation results. Each column represents a local trait. If a row 
corresponds to a nonlocal trait that is mediated by the local trait, then the cell is colored dark green; 
otherwise, the cell is light green. The last annotation column and the annotation row above the heatmap both 
designate local traits with dark blue and nonlocal traits with light blue.
